Analytics Without Parameter Tuning Considered Harmful?

### 2015-03-14 tuningpredictor.tex
Updates from ShareLaTeX

Firstly, we are {\em not} saying that {\em all} learning requires
tuning. For example, when proposing fixes to software, Weimar et al. use a genetic learner
to propose patches~\cite{Weimer:2009}. Consider one performance criteria for that work; i.e. 
that   that that method can find and fix known bugs. Note that this is a {\em competency criteria}
which does not include the phrase  {\em better than}. Such performance criteria do
not requiring tuning.  However, once {\em better than} enters the performance criteria
(as done in \cite{lessmann2008benchmarking,hall11,me07b,bell2013limited,rahman2013how,me02k,moser2008comparative,zimmermann2007predicting,herzig2013predicting})
then this becomes a race between competing methods (or attributes).  In such a race, it is unfair
to hobble one competitor with poor tunings.

@@

\section{Motivating Example}\label{sec:eg}

@@

One way to generate a  defect predictor from
static code is to use 
linear  regression.
This is a standard statistical
method that fits a straight line to a set of points. The line
offers a set of predicted values.
If the points are somewhat
scattered, then a single regression line cannot pass through all points and the distance from these predicted values to the actual
values is a measure of the error associated with that line.
Linear regression search for a line that minimizes that
error and maximizes the correlation\footnote{
$Correlation$
measures how closely two variables co-vary.  It ranges from   from -1
(perfect negative correlation) through 0 (no correlation) to +1
 (perfect positive correlation).
Let  $a_i$ and $p_i$ denote some actual and predicted values respectively; and  $n$ and $\overline{x}$ denote
 the number of observations and the mean of the $n$ observations, 
respectively. Then:
$S_{PA}=(\sum_i (p_i - \overline{p})(a_i -
  \overline{a}))/(n-1)$ and
$S_{p}=(\sum_i (p_i - \overline{p})^2)(n-1)$
and 
$S_{a}=(\sum_i (a_i - \overline{a})^2)/(n-1)$ and
correlation is  $c=S_{PA}/\sqrt{S_pS_a}$.}  denoted as $c$.

@@

{\scriptsize \[\begin{array}{l|l|ll}
\mathit{measures} & d= \mathit{\#defects} & \mathit{correlation}\\\hline
\mathit{LOC}   &d_i= 0.0164 +0.0114\mathit{LOC}\ & c_1 = 0.65\\\hline
\mathid{Halstead & d_2= 0.231 + 0.00344N     \\    
           & +  0.0009V   - 0.185L \\
           & - 0.0343D      - 0.00541I \\
           & + 0.000019E + 0.711B \\ 
            &   - 0.00047T  & c_2=-0.36 
            \end{array}
\end{eqnarray}}

\noindent
Enter tuning. Suppose the defect predictors $d_1$ and $d_2$  learned from LOC or Halstead
are used to to call an inspection
team to check for errors in certain parts of the code using the rule
\begin{equation}\label{eq:yesno}\scriptsize
\mathit{inspect}= \left\{
\begin{array}{ll}
d_i \ge T \rightarrow \mathit{Yes}\\
d_i <   T \rightarrow \mathit{No}
\end{array}
\end{equation}
\fig{pd1} shows the effects of tuning the $T$ variable. Not surprisingly,
at $T=0$, all modules get inspected so the false alarm rate is very high. To reduce that
problem, we can increase $T$. \fig{pd1} reports that the false alarm rate falls below
20\% at $T=0.45$ (for Halstead).

@@

\subsection{Differential Evolution Algorithm}
Differential Evolution (DE)\cite{storn1997differential} is a stochastic search algorithm that 
optimizes a problem by iteratively trying to improve a  population of candidate solutions with 
regard to a given quality measurement. Such method makes no assumptions about the 
problem being optimized and has already been used as a parameter tuner
\cite{omran2005differential, chiha2012tuning}.

@@

The experiment aims at investigate whether tuning helps learners improve performance in 
terms of accuracy measurement. We choose compare WHERE-based learner with and 
without tuning parameters and two {\it significantly different} learners Random Forest and 
CART according to \cite{lessmann2008benchmarking}. As mentioned above, we'd like to see 
whether tuning will help change the rank of CART and make it comparable with Random 
Forest.

To evaluate accuracy performance of learners, several measurements are proposed, like 
probability of  detection({\it pd}) and probability of false alarm ({\it pf})\cite{menzies2007data}, 
the area under the receiver operating characteristics curve ({\it AUC})
\cite{lessmann2008benchmarking}, and precision\cite{zhang2007comments}. In this work, we 
expect learners should identify as many defective modules as possible while avoiding false 
alarm. Therefore, learners are evaluated by both of {\it pd} and {\it pf} simaltanieously. A single 
measure, G-measure, defined as the harmonic mean of {\it pd} and $1-{\it pf}$  is used. The 
G-measure value is between 0 and 1. The higher, the better.
\begin{equation}
G = \frac{2*(1-pf)*pd}{1-pf+pd}
\end{equation}

In this experiment, we use three different portions of one project data set for training, tuning 
and testing process. In contrast to hold out way used in \cite{lessmann2008benchmarking, 
menzies2007data}, we separate the data sets in order. Since learners are designed to predict 
defects in future projects,  any randomly data set selection without taking into the time series 
will not sufficient to evaluate the performance of predicting future. To the most, that is good to 
evaluate the accuracy of classification but not predicting future. Since we have 10 different 
project data, each of which contains least 3 evolutionary versions. We use the following policy 
to select the data: in each project, we only use the last three data files for experiment. 
Specifically, the $nth$,  $(n-1)th$, $(n-2)th$ versions of project data are selected for testing, 
tuning and training learners, respectively. This will make sure that we don't use the future 
project data to train learners and predict previous project.

@@

Besides the Where-based learner implemented by ourself,  we use the CART and Random Forest modules 
from scikit-learn \cite{scikit-learn} for this experiment. The parameters associated with different learners are listed in Fig.\ref{fig:parameters}. (\textbf{NEED to elaborate that there're different versions of CART, what's the point to do this experiment}). For each data set, run CART, naive Where-based learner and Random Forests with corresponding default values. Then using DE to tune corresponding parameters for CART and Where-based Learner,  and run them again with the optimal parameters from tuning process to test the performance. 
This study ranks learners using the Scott-Knott procedure recommended by Mittas \& Angelis in their 2013
IEEE TSE paper~\cite{mittas2013ranking}.  This method sorts a list of $l$ treatments with $ls$ measurements by their median score. It then splits $l$ into sub-lists $m,n$ in order to maximize the expected value of differences  in the observed performances before and after divisions.

@@

@@

\bibliographystyle{unsrt}
\bibliography{tuningpredictor}

%%%%%%%%%%%%%%%% list of parameters%%%%%%%%%%%%%%%%%%%%%
\renewcommand\arraystretch{1.2}
\begin{figure*}[t!]
\renewcommand{\baselinestretch}{0.8}
\scriptsize
  \centering
\begin{tabular}{|c|c|c|c|l|}
\cline{1-5}
\begin{tabular}[c]{@{}c@{}}Learner \\ Name\end{tabular} & Parameters & Default &\begin{tabular}[c]{@{}c@{}}Tuning\\ Range\end{tabular}& 
\multicolumn{1}{c|}{Description} \\ \hline
\multirow{8}{*}{\begin{tabular}[c]{@{}c@{}}Where-based\\ Learner\end{tabular}} 
& threshold & 0.5 &[0.01,1]& The value to determine defective or not .\\ \cline{2-5} 
& infogain & 0.33 &[0.01,1]& The percentage of features to consider  for the best 
split to build CART tree\footnote{Since the Where-based learner will build two trees, the first 
one is for clustering and the second one is building prediction model. we explicitly call Where-
clustering tree and CART tree, respectively}. \\ \cline{2-5} 
& min\_sample\_size & 4& [1,10]& The minimum number of samples required to be a leaf for 
CART tree. \\ \cline{2-5} 
& min\_Size & 0.5 &[0.01,1]& \begin{tabular}[c]{@{}l@{}}The value to determine the minimum 
number of samples to be a Where-clustering tree \\ based on  ${n\_samples}^ {min\_Size}$.
\end{tabular} \\ \cline{2-5} 
    & wriggle & 0.2 &[0.01, 1] & The threshold to determine which branch in  Where tree to be pruned\\ \cline{2-5}
& depthMin & 2 & [1,6]&The minimum depth of the tree below which no pruning for Where-
clustering tree. \\ \cline{2-5} 
& depthMax & 10 &[1,20]& The maximum depth of the Where-clustering tree. \\ \cline{2-5} 
& wherePrune & False &T/F& Whether or not to prune the Where-clustering tree. \\ \cline{2-5}
& treePrune & True &T/F& Whether or not to prune the classification tree built by CART. \\ \cline{2-5} 
\hline
\multirow{4}{*}{CART} & threshold & 0.5 &[0,1]& The value to determine defective or not. \\ \cline{2-5} 
& max\_feature & None &[0.01,1]& The number of features to consider when looking for the best 
split. \\ \cline{2-5}

@@

We write to suggest that all those conclusions may be wrong or, at the very most,
correct within a very limited context. At the very least, the conclusions
of those papers (and literally dozens of similar papers, include severl
of our own~\cite{me02e,me02k,me03a,me06d,me07b}) need to be revisited
in light of this paper's results on the effects of tuning on data miners.

>>>>

Firstly, we are {\em not} saying that {\em all} learning requires
tuning. For example, when proposing fixes to software, Weimar et al. use a genetic learner
to propose patches~\cite{Weimer:2009}. Consider one performance criteria for that work; i.e. 
that   that that method can find and fix known bugs. Note that this is a {\em competency criteria}
which does not include the phrase  ``{\em better than}''. Such performance criteria do
not requiring tuning.  However, once {\em better than} enters the performance criteria
(as done in \cite{lessmann2008benchmarking,hall11,me07b,bell2013limited,rahman2013how,me02k,moser2008comparative,zimmermann2007predicting,herzig2013predicting})
then this becomes a race between competing methods (or attributes).  In such a race, it is unfair
to hobble one competitor with poor tunings.

@@

\section{Motivating Example}\label{sect:eg}

@@

One way to generate a  defect predictor from
static code is to use 
linear  regression.
This is a standard statistical
method that fits a straight line to a set of points. The line
offers a set of predicted values.
If the points are somewhat
scattered, then a single regression line cannot pass through all points and the distance from these predicted values to the actual
values is a measure of the error associated with that line.
Linear regression search for a line that minimizes that
error and maximizes the correlation\footnote{
$Correlation$
measures how closely two variables co-vary.  It ranges from   from -1
(perfect negative correlation) through 0 (no correlation) to +1
 (perfect positive correlation).
Let  $a_i$ and $p_i$ denote some actual and predicted values respectively; and  $n$ and $\overline{x}$ denote
 the number of observations and the mean of the $n$ observations, 
respectively. Then:
$S_{\mathit{PA}}=(\sum_i (p_i - \overline{p})(a_i -
  \overline{a}))/(n-1)$ and
$S_{p}=(\sum_i (p_i - \overline{p})^2)(n-1)$
and 
$S_{a}=(\sum_i (a_i - \overline{a})^2)/(n-1)$ and
correlation is  $c=S_{\mathit{PA}}/\sqrt{S_pS_a}$.}  denoted as $c$.

@@

{\scriptsize \[
\begin{array}{l|l|ll}
\mathit{measures} & d= \mathit{\#defects} & \mathit{correlation}\\\hline
\mathit{LOC}   &d_i= 0.0164 +0.0114\mathit{LOC}\ & c_1 = 0.65\\\hline
\mathit{Halstead} & d_2= 0.231 + 0.00344N     \\    
                 & +  0.0009V   - 0.185L \\
                 & - 0.0343D      - 0.00541I \\
                 & + 0.000019E + 0.711B \\ 
                 &   - 0.00047T  & c_2=-0.36  
\end{array}
\]
}

\noindent
Enter tuning. Suppose the defect predictors $d_1$ and $d_2$  learned from LOC or Halstead
are used to to call an inspection
team to check for errors in certain parts of the code using the rule
\begin{equation}\label{eq:yesno}\scriptsize
\mathit{inspect}= \left\{
\begin{array}{ll}
d_i \ge T \rightarrow \mathit{Yes}\\
d_i <   T \rightarrow \mathit{No} 
\end{array}\right.
\end{equation}
\fig{pd1} shows the effects of tuning the $T$ variable. Not surprisingly,
at $T=0$, all modules get inspected so the false alarm rate is very high. To reduce that
problem, we can increase $T$. \fig{pd1} reports that the false alarm rate falls below
20\% at $T=0.45$ (for Halstead).

@@

The psuedocde for differential evolution is shown in Algorithm \ref{alg:DE}

for 
Differential Evolution (DE)\cite{storn1997differential} is a stochastic search algorithm that 
optimizes a problem by iteratively trying to improve a  population of candidate solutions with 
regard to a given quality measurement. Such method makes no assumptions about the 
problem being optimized and has already been used as a parameter tuner
\cite{omran2005differential, chiha2012tuning}.

@@

The experiment aims at investigating whether tuning helps learners improve performance in 
terms of accuracy measurement. We choose compare WHERE-based learner with and 
without tuning parameters and two {\it significantly different} learners Random Forest and 
CART according to \cite{lessmann2008benchmarking}. As mentioned above, we'd like to see 
whether tuning will help change the rank of CART and make it comparable with Random 
Forest.

To evaluate accuracy performance of learners, several measurements are proposed to use, like 
probability of  detection({\it pd}) and probability of false alarm ({\it pf})\cite{menzies2007data}, 
the area under the receiver operating characteristics curve ({\it AUC})
\cite{lessmann2008benchmarking}, precision and F measure\cite{zhang2007comments}. In this work, we 
expect learners should identify as many defective modules as possible while avoiding false 
alarm. Therefore, learners are evaluated by both of {\it pd} and {\it pf} simaltanieously. A single 
measure, G-measure, defined as the harmonic mean of {\it pd} and $1-{\it pf}$  is used. The 
G-measure value is between 0 and 1. The higher, the better.
\begin{equation}
G = \frac{2*(1-pf)*pd}{1-pf+pd}
\end{equation}

In this experiment, we divide one project data set into training, tuning 
and testing datasets. In contrast to hold out way used in \cite{lessmann2008benchmarking, 
menzies2007data}, we separate the data sets in order. Since learners are designed to predict 
defects in future projects,  any randomly data set selection without taking into the time series 
will not sufficient to evaluate the performance of predicting future. Since we have data from 
10 different projects, each of which contains at least 3 evolutionary versions. We use the following policy 
to select the data for experiments: for each project, we use every three chronological data files 
to do testing, tuning and training in one experiment. For example, in ant project,
we have versions of $ant1.3$, $1.4$, $1.5$, $1.6$, and $1.7$, which will generate data sets called
 $antV0$, $antV1$ and $antV2$ for three experiments.

@@

Besides the Where-based learner implemented by ourselves,  we use the CART and Random Forest modules 
from scikit-learn \cite{scikit-learn} to do a comparative study. The parameters associated with different learners are listed in Fig.\ref{fig:parameters}. For each data set, run CART, naive Where-based learner and Random Forests with corresponding default values. Then using DE to tune corresponding parameters for CART and Where-based Learner,  and run them again with the optimal parameters from tuning process to test the performance. 
% This study ranks learners using the Scott-Knott procedure recommended by Mittas \& Angelis in their 2013
% IEEE TSE paper~\cite{mittas2013ranking}.  This method sorts a list of $l$ treatments with $ls$ measurements by their median score. It then splits $l$ into sub-lists $m,n$ in order to maximize the expected value of differences  in the observed performances before and after divisions.

@@

\begin{figure}[!t]

\renewcommand{\baselinestretch}{0.8}
\scriptsize
\centering
  \begin{tabular}{r@{~}|r@{~}rr|r@{~}rr|r@{~}rr}
    & \multicolumn{3}{c|}{WHERE}&\multicolumn{3}{c|}{CART}& \multicolumn{3}{c}{Random Forest}\\\cline{2-10} 
    Goal & tuned & naive &$\frac{\mathit{tuned}}{\mathit{naive}}$& tuned & naive &$\frac{\mathit{tuned}}{\mathit{naive}}$& tuned & naive&$\frac{\mathit{tuned}}{\mathit{naive}}$\\\hline
pd&71.4&1.2&60&4.4&0.1&44&7.1&0.2&36\\ 
prec&82.8&1.4&59&3.9&0.1&39&7.4&0.2&37\\
f&86.1&1.2&72&3.3&0.1&33&6.7&0.2&34
  \end{tabular}
  \caption{For one train/test set pair (ANT),  time (in seconds) spent tuning,
  for different goals, or just running the ``off-the-shelf'' learner.}
\end{figure}

\begin{figure}[!t]

\renewcommand{\baselinestretch}{0.8}
\scriptsize
\centering
\begin{tabular}{r|rrr}
goal &	WHERE	&CART&	Random Forest\\\hline
f	&15	&13	&9\\ 
pd	&15	&16	&8\\ 
prec&	14	&14	&8
\end{tabular}
\caption{How much slower is tuning vs just running the default
settings of the learner? E.g. bottom right, tuning to increase
precision with Random Forests is eight times slower than
just running Random Forests.}
\end{figure}

@@

\vspace*{0.5mm}

 \scriptsize
\bibliographystyle{unsrt}
\bibliography{tuningpredictor}

%%%%%%%%%%%%%%%% list of parameters%%%%%%%%%%%%%%%%%%%%%
\renewcommand\arraystretch{1.2}
\begin{figure*}[t!]
\renewcommand{\baselinestretch}{0.8}
\scriptsize
  \centering
\begin{tabular}{|c|c|c|c|l|}
\cline{1-5}
\begin{tabular}[c]{@{}c@{}}Learner \\ Name\end{tabular} & Parameters & Default &\begin{tabular}[c]{@{}c@{}}Tuning\\ Range\end{tabular}& 
\multicolumn{1}{c|}{Description} \\ \hline
\multirow{8}{*}{\begin{tabular}[c]{@{}c@{}}Where-based\\ Learner\end{tabular}} 
& threshold & 0.5 &[0.01,1]& The value to determine defective or not .\\ \cline{2-5} 
& infogain & 0.33 &[0.01,1]& The percentage of features to consider  for the best 
split to build CART tree\footnote{Since the Where-based learner will build two trees, the first 
one is for clustering and the second one is building prediction model. we explicitly call Where-
clustering tree and CART tree, respectively}. \\ \cline{2-5} 
& min\_sample\_size & 4& [1,10]& The minimum number of samples required to be a leaf for 
CART tree. \\ \cline{2-5} 
& min\_Size & 0.5 &[0.01,1]& \begin{tabular}[c]{@{}l@{}}The value to determine the minimum 
number of samples to be an internal node of the Where-clustering tree \\ based on  ${n\_samples}^ {min\_Size}$.
\end{tabular} \\ \cline{2-5} 
    & wriggle & 0.2 &[0.01, 1] & The threshold to determine which branch in  Where tree to be pruned\\ \cline{2-5}
& depthMin & 2 & [1,6]&The minimum depth of the tree below which no pruning for Where-
clustering tree. \\ \cline{2-5} 
& depthMax & 10 &[1,20]& The maximum depth of the Where-clustering tree. \\ \cline{2-5} 
& wherePrune & False &T/F& Whether or not to prune the Where-clustering tree. \\ \cline{2-5}
& treePrune & True &T/F& Whether or not to prune the classification tree built by CART. \\ \cline{2-5} 
\hline
\multirow{4}{*}{CART} & threshold & 0.5 &[0,1]& The value to determine defective or not. \\ \cline{2-5} 
& max\_feature & None &[0.01,1]& The number of features to consider when looking for the best 
split. \\ \cline{2-5}

@@

We write to suggest that all those conclusions may be wrong or, at the very most,
correct within a very limited context. At the very least, the conclusions
of those papers (and literally dozens of similar papers, include several
of our own~\cite{me02e,me02k,me03a,me06d,me07b}) need to be revisited
in light of this paper's results on the effects of tuning on data miners.
